Manuscript results: Forecasting influenza using ARIMA and ARIMAX

### 2025-06-20 manuscript_results.Rmd
code review

Here we will be computing and plotting mean WIS improvement by epidemiological weeks comparing the best model and the basiline model.

```{r}
# Create a list of data frames for the four weeks
weeks <- 1:4
summed_data_list <- list()

for (week in weeks) {
  # Merge the data frames for the current week
  data__ <- merge(get(paste0("filtered_df_W", week, "_NoLg")),
                  get(paste0("filtered_df_W", week)))

  # Sum values by Julian date across all states
  summed_data <- data__ %>%
    group_by(Julian_date) %>%
    summarize(
      AUTO_AR = mean(AUTO_AR, na.rm = TRUE),  # Sum across all states
      AUTO_AVG_LB = mean(AUTO_AVG_LB, na.rm = TRUE)           # Sum across all states
    ) %>%
    # Calculate the difference ratio
    mutate(Difference =  ((AUTO_AVG_LB / AUTO_AR)-1)*100,
           Week = paste0("Week ", week))

  # Store the summarized data in the list
  summed_data_list[[week]] <- summed_data
}

# Combine all weeks into a single data frame
summed_data_all <- bind_rows(summed_data_list)

# Plot the difference using ggplot2
model_improv<-ggplot(summed_data_all, aes(x = Julian_date, y = (Difference), color = Week)) +
  geom_point(size = 2.5) +
  geom_hline(yintercept = 0, color = "black", linetype = "dashed", size = 0.8) +
  theme_minimal() +
  labs(
    title = "A) Weekly mean WIS differences across all states",
    subtitle = "AUTO_AVG_LB compared to baseline",
    x = "",
    y = "Mean difference"
  ) +
  scale_y_continuous(
    limits = c(-100, 100),
    breaks = seq(-100, 100, by = 20)
  ) +
  scale_x_date(
    date_breaks = "1 month",
    date_labels = "%b %y"
  ) +
scale_color_manual(
  values = c(
    "Week 1" = "#4575B4",  # Blue
    "Week 2" = "#91BFDB",  # Light blue
    "Week 3" = "#E89C9C",  # Light red
    "Week 4" = "#D73027"   # Red
  )
) +
  theme(
    plot.title = element_text(size = 18, face = "bold"),
    plot.subtitle = element_text(size = 14),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

@@

total_hospitalizations_by_week <- AUTO_ADJACENT_WEEK1 %>%
  filter(epiweek >= 40 | epiweek <= 20) %>%  # Apply filter AFTER epiweek exists
  group_by(target_end_date) %>%
  summarise(mean_cases = sum(cases, na.rm = TRUE))

hospital_plot <-ggplot(total_hospitalizations_by_week, aes(x = target_end_date, y = mean_cases)) +
  geom_bar(stat = "identity",fill = "steelblue") +
  labs(title = "B) Total influenza hospitalizations on the 48 states",
       x = "",
       y = "Influenza hospitalizations") +
  scale_x_date(
    date_breaks = "1 month",
    date_labels = "%b %y"
  ) +
  theme_minimal()+
  theme(plot.title = element_text(size = 18, face = "bold"),
    plot.subtitle = element_text(size = 14),
    axis.text.x = element_text(angle = 45, hjust = 1))

@@

# Combine the plots vertically
combined_plot <-  model_improv /hospital_plot

# Display (optional)
combined_plot

>>>>

Here we will be computing and plotting mean WIS improvement by epidemiological weeks comparing the best model and the baseline model.

```{r}
# Create a list of data frames for the four weeks
weeks <- 1:4
results_by_epiweek_list <- list()

for (week in weeks) {
  # Merge the data frames for the current week
  combined_dataframes_ <- merge(get(paste0("filtered_df_W", week, "_NoLg")),
                  get(paste0("filtered_df_W", week)))

  # Sum values by Julian date across all states
  mean_improvement_ <- combined_dataframes_ %>%
    group_by(Julian_date) %>%
    summarize(
      AUTO_AR = mean(AUTO_AR, na.rm = TRUE),  # Sum across all states
      AUTO_AVG_LB = mean(AUTO_AVG_LB, na.rm = TRUE)           # Sum across all states
    ) %>%
    # Calculate the difference ratio in percentage
    mutate(Difference =  ((AUTO_AVG_LB / AUTO_AR)-1)*100,
           Week = paste0("Week ", week))

  # Store the mean improvement by julian date
  results_by_epiweek_list[[week]] <- mean_improvement_
}

# Combine all weeks into a single data frame
results_by_epiweek_all <- bind_rows(results_by_epiweek_list)

# Plot the results
model_improv<-ggplot(results_by_epiweek_all, aes(x = Julian_date, y = (Difference), color = Week)) +
  geom_point(size = 2.5) +
  geom_hline(yintercept = 0, color = "black", linetype = "dashed", size = 0.8) +
  theme_minimal() +
  labs(
    title = "A) Weekly mean WIS differences across all states",
    subtitle = "AUTO_AVG_LB compared to baseline",
    x = "",
    y = "Mean difference"
  ) +
  scale_y_continuous(
    limits = c(-100, 100),
    breaks = seq(-100, 100, by = 20)
  ) +
  scale_x_date(
    date_breaks = "1 month",
    date_labels = "%b %y"
  ) +
scale_color_manual(
  values = c(
    "Week 1" = "#4575B4",  # Blue
    "Week 2" = "#91BFDB",  # Light blue
    "Week 3" = "#E89C9C",  # Light red
    "Week 4" = "#D73027"   # Red
  )
) +
  theme(
    plot.title = element_text(size = 18, face = "bold"),
    plot.subtitle = element_text(size = 14),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

@@

# get total hospitalizations for the 48 states
total_hospitalizations_by_week <- AUTO_ADJACENT_WEEK1 %>%
  filter(epiweek >= 40 | epiweek <= 20) %>%  
  group_by(target_end_date) %>%
  summarise(mean_cases = sum(cases, na.rm = TRUE))

# plotting results
hospital_plot <-ggplot(total_hospitalizations_by_week, aes(x = target_end_date, y = mean_cases)) +
  geom_bar(stat = "identity",fill = "steelblue") +
  labs(title = "B) Total influenza hospitalizations on the 48 states",
       x = "",
       y = "Influenza hospitalizations") +
  scale_x_date(
    date_breaks = "1 month",
    date_labels = "%b %y"
  ) +
  theme_minimal()+
  theme(plot.title = element_text(size = 18, face = "bold"),
    plot.subtitle = element_text(size = 14),
    axis.text.x = element_text(angle = 45, hjust = 1))

@@

```{r}
# Combining results in a single plot
combined_plot <-  model_improv /hospital_plot 
combined_plot
